\documentclass[10pt]{book}

\usepackage{bm,amsmath}
\usepackage{makeidx}
\usepackage{color}
\usepackage{natbib}
\usepackage{mfpic}
\usepackage{pxfonts}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{verbatim} 
\usepackage{authblk}
\usepackage{tikz}
\usepackage[framemethod=tikz]{mdframed}

\usepackage{titlesec}
\titleformat*{\section}{\LARGE\bfseries}
\titleformat*{\subsection}{\Large\bfseries}
\titleformat*{\subsubsection}{\large\bfseries}
\titleformat*{\paragraph}{\large\bfseries}
\titleformat*{\subparagraph}{\large\bfseries}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{minitoc} 
\renewcommand{\mtcfont}{\small\sffamily} 
\renewcommand{\mtcSfont}{\small\sffamily} 
\usepackage[Lenny]{./styles/fncychap}

%\evensidemargin -.2in
%\oddsidemargin .2in
\evensidemargin -.3in
\oddsidemargin  0.0in

\topmargin -0.50in
\textheight 9.0in
\renewcommand{\baselinestretch}{1.0}
\textwidth 6.5in


\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\leftmark}
\rhead{Section \thesection}
\lfoot{{\sc CVMix theory and numerics}}
\cfoot{\today}
\rfoot{Page \thepage}
\renewcommand{\footrulewidth}{1pt}
\usepackage{kpfonts}
\fancyhead[LE,LO]{\textsc{\textbf{\nouppercase{\leftmark}}}}

\usepackage{titlesec,titletoc} 
\titleclass{\part}{top}
\titleformat{\part}
{\centering\normalfont\Huge\bfseries}{}{0pt}{}

\usepackage{./styles/Griffies_style}

\usepackage{hyperref}  % must be used last in the list of packages 
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue} 


\setcounter{tocdepth}{1}
\setcounter{minitocdepth}{3}
\setcounter{secnumdepth}{3}

\makeindex 
\citeindextrue

\title{\sc Theory and Numerics of the Community 
         \\ Ocean Vertical Mixing (CVMix) Project}

\date{\today}

\author[$\star$]{Stephen M. Grif\/f\/ies \thanks{Stephen.Griffies@noaa.gov}}
\affil[$\star$]{NOAA/Geophysical Fluid Dynamics Laboratory, Princeton USA}

\author[$\dagger$]{Michael Levy \thanks{MLevy@ucar.edu}}
\affil[$\dagger$]{National Center for Atmospheric Research, Boulder USA}

\author[$\star$]{Alistair J. Adcroft \thanks{Alistair.Adcroft@noaa.gov}}
\author[$\dagger$]{Gokhan Danabasoglu \thanks{Gokhan@ucar.edu}}
\author[$\star$]{Robert W. Hallberg \thanks{Robert.Hallberg@noaa.gov}}

\author[$\triangle$]{Doub Jacobsen \thanks{Jacobsen.douglas@gmail.com}}
\affil[$\triangle$]{Los Alamos National Laboratory, Los Alamos USA}

\author[$\dagger$]{William Large \thanks{Wily@ucar.edu}}
\author[$\triangle$]{Todd Ringler \thanks{Ringler@lanl.gov}}

\begin{document}

\pagenumbering{roman}
\maketitle 
\thispagestyle{empty}
%\newpage 

\begin{center}
{\scshape \Large The CVMix Project} 
\end{center}

{\sc Community Ocean Vertical Mixing (CVMix)} is a software package
that aims to provide transparent, robust, flexible, well documented,
shared Fortran source code for use in parameterizing vertical mixing
processes in numerical ocean models.  The project is focused on
developing software for a consensus of first-order closures that
return a vertical diffusivity, viscosity, and possibly a non-local
transport, with each quantity dependent on the tracer or velocity
being mixed.  CVMix modules are written as kernals designed for use in
a variety of Fortran ocean model codes such as MPAS-ocean, MOM, and
POP.  CVMix modules use MKS units and expect the same for input and
output.  Code development occurs within a community of scientists and
engineers who make use of CVMix modules for a variety of ocean codes.
When mature, CVMix modules will be freely distributed to the open
source community under GPLv2 using an open source methodology.

As of August 2013, the CVMix code includes a suite of mixing schemes,
including KPP. Significant refinement to the code will follow as
testing within MPAS, MOM, and POP continues. 


\vspace{2cm}

\noindent
This document is freely distributed and should be referenced as the
following. \vspace{.25cm}

\noindent
{\bf 
{\scshape{Theory and Numerics of the Community Ocean Vertical
    Mixing (CVMix) Project}}
\\
S.M. Grif\/f\/ies, M. Levy, A.J. Adcroft, G. Danabasoglu,
R.W. Hallberg,  D. Jacobsen, W. Large, T. Ringler
\\
Technical Report
\\
Draft from \today 
\\
\pageref{LastPage} + v pages\\
}
\vspace{.4cm}



\noindent
This document was prepared using \LaTeX \hspace{.1cm} as described by
\cite{Latex} and \cite{LatexCompanion}.





\newpage 

\dominitoc
\tableofcontents
\listoffigures

\newpage 
\pagenumbering{arabic}

\input{cvmix_intro}
\input{cvmix_elements}
\input{cvmix_background}
\input{cvmix_shear}
\input{cvmix_tidal}
\input{cvmix_ddiff}
\input{cvmix_kpp}
%\input{cvmix_convection}
%\input{cvmix_dissipation}


%%%%%%%%%%%%%%%%%%%%%      References     %%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 

\bibliographystyle{./references/elsarticle-harv.bst}
\addcontentsline{toc}{chapter}{\scshape Bibliography} 
\markboth{\scshape BIBLIOGRAPHY}{\scshape BIBLIOGRAPHY}
\gdef\rightmark{\scshape BIBLIOGRAPHY}
\bibliography{./references/Griffies_refs}
%\bibliography{/Users/smg/Documents/TEX/bibtex/Griffies_refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\printindex

\end{document}








